---
title: "lab7"
author: "R. Spellenberg"
date: "2023-02-23"
output: 
  html_document:
    code_folding: hide
---

```{r setup, echo = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, results = FALSE)
library(tidyverse)
library(here)
library(broom)

### Part One: Wrangling and Visualizing Spatial Data
library(sf) ### simple features
library(tmap)

### Part Two: Spatial Analysis and Kriging
library(gstat)
library(stars)

```

# Part One

### Read in CA county outlines 
```{r}
ca_counties_sf <- read_sf(here("data/ca_counties/CA_counties_TIGER2016.shp"))

### geometry column is sticky unless you intentionally drop it
ca_subset_sf <- ca_counties_sf %>% 
  janitor::clean_names() %>% 
  select(county_name = name, land_area = aland)

### drops geometry column
ca_subset_df <- ca_subset_sf %>% 
  as.data.frame() %>% 
  select(-geometry)
```

### Check and set the CRS (Coordinate Reference System)
```{r}
ca_subset_sf %>% st_crs() ### or st_crs(ca_subset_sf)

ggplot() + 
  geom_sf(data = ca_subset_sf, aes(fill = land_area),
          color = 'white', size = 0.1) +
  scale_fill_gradientn(colors = c('cyan', 'blue', 'purple')) +
  theme_void()
```

### Load another dataset on an invasive species
```{r}
sesbania_sf <- read_sf(here("data/red_sesbania/ds80.shp")) %>% 
  janitor::clean_names()

sesbania_sf %>% st_crs()

sesbania_3857_sf <- sesbania_sf %>% 
  st_transform(3857)

```

### Plot the two together
```{r}
### the order you place the layers builds on top of each other (first placed is the base layer)

ggplot() +
  geom_sf(data = ca_subset_sf, fill = 'gray70') +
  geom_sf(data = sesbania_3857_sf, size = 1, color = 'red') +
  theme_void()
```

```{r}
ca_sesb_sf <- ca_subset_sf %>% 
  st_join(sesbania_3857_sf)

sesb_counts_sf <- ca_sesb_sf %>% 
  group_by(county_name) %>% 
  summarize(n_records = sum(!is.na(id)))
```

### Plot the results

"Choropleth" map
```{r}
ggplot(data = sesb_counts_sf %>% 
         mutate(n_records = ifelse(n_records == 0, NA, n_records))) +
  geom_sf(aes(fill = n_records), color = 'gray20', size = 0.1) + 
  geom_sf(data = sesbania_3857_sf, color = 'midnightblue') +
  scale_fill_gradientn(colors = c('lightyellow', 'orange', 'red'), na.value = 'lightgray') + theme_void()
```

### Plot just the points in Solano County
```{r}
solano_sesb_sf <- sesbania_3857_sf %>% 
  filter(county == 'Solano')

solano_sf <- ca_subset_sf %>% 
  filter(county_name == 'Solano')

ggplot() +
  geom_sf(data = solano_sf) +
  geom_sf(data = solano_sesb_sf, color = 'midnightblue') +
  theme_void()
```

### Make an interactive map
```{r}
### set tmap interactive mode
tmap_mode('view') ### can use 'plot' to make static map

tm_shape(ca_subset_sf) +
  tm_fill("land_area", palette = 'BuGn') +
  tm_shape(sesbania_3857_sf) +
  tm_dots()

### look up vignette on using tmap
```

Layer Sources
- naturalearthdata.com (rpackage - 'rnaturalearth')
?rnaturalearth in the console will tell you about the arguments of the package

